The yaq Project: a Protocol for Scientific Instrumentation

### 2022-04-29 paper.tex
skeleton

@@

@@

@@

A few ``short stories'' of particular instruments utilizing yaq.

\subsubsection{yaqc}

Lightweight scriptable.

\subsubsection{yaqc-qtpy}

Capitalizes heavily on traits

\subsubsection{Stahl}

Self contained ecosystem on a raspberry pi.

Dedicated client application.

Exceedingly simple.

(reference publication)

\subsubsection{Landis}

....

\subsubsection{Goldsmith}

Mixing with other hand built or commercially derived software.

\subsubsection{Bluesky}

Bluesky is an extensive open source control layer \cite{AllanDanielB2019a}.
Bridging to larger open source control layers.

>>>>

\subsection{TCP}

\subsection{Avro RPC}

@@

\subsection{Configuration and State}

@@

\subsection{Traits}

Graphic of traits?

Examples of daemons for core traits.

@@

yaq's RPC design invites the creation of multiple client libraries specialized for specific purposes.
Again, daemons can communicate simultaniously with multiple clients.
For a yaq instrument it is normal to have multiple clients running simultaniously, each focused on a particular purpose.
This modularity speeds development and increases overall flexibility of the system.
We have created a few client libraries which we will discuss here.

yaqc [CITE] is a very basic Python client for yaq.
yaqc is totally generic and aims to support any conceivable yaq RPC.
This library is easy to use directly in Python scripts or within the REPL, but it's also very useful as a dependency when building more specialized Python clients.

yaqc-qtpy [CITE] is a specialty Qt-based [CITE] GUI client.
yaqc-qtpy provides a special QClient object which uses Qt threading primatives to provide performant and safe access to yaq from a Qt environment.
In addition, yaqc-qtpy provides an entry point which automatically builds a GUI from arbitrary protocols using traits.
This is an invaluable tool which provides a ``free'' graphical user interface to any daemon.

Bluesky is an extensive open source control layer developed collaboratively across several user facilities \cite{AllanDanielB2019a}.

Bridging to larger open source control layers.
PyMoDaq \cite{WeberSebastien2021a}

\section{Case Studies}

Introductory paragraph.

The Landis Group at UW-Madison created WiQK, a quench kinetics reactor.
Reactions can be orchestrated with a simple script using Python's built-in time library and yaqc.
This has worked for a few years as developers continue to modify their reactor.

The Stahl Group at UW-Madison created a custom reactor which monitors gasses being produced or consumed in the reaction head-space.  \cite{SalazarChaseA2021a}
From an automation control perspective, this reactor incorporates a collection of sensors and a single heating process value under computer control
This reactor was made as a self-contained unit with a Raspberry Pi [CITE] as the computer.
Users interact with a custom Qt-based GUI to use the reactor.
yaq daemons run locally along-side the lightweight Qt application making the entire system simple to run.

The Wright Group needs to orchestrate a large variety of hardware in multidimensional scans for their complex spectroscopy experiments.
We have built out a toolchain incorporating yaq into Bluesky to take advantage of the sophisticated orchestration that Bluesky supports.
We are excited about Bluesky moving forward.
